---
title: "Machine Learning Assigment"
author: "Pablo Rojo"
date: "24 de agosto de 2014"
output: html_document
---

#Introduction
In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset). 

#Data Exploration and Cleaning
First we load the data provided for training:
```{r}
data <- read.csv("pml-training.csv")
```

Since cross validation will be used, the data is split in training and testing sets:
```{r}
library(caret)
set.seed(2235)
inTrain <- createDataPartition( y=data$classe,
                                p=0.75, list=FALSE)
training <- data[inTrain,]
testing <- data[-inTrain,]
```

A first look at the data using:

```{r, eval=FALSE}
summary(training)
```

show:

- 160 variables
- Multiple columns have mostly NA.
- The first 7 variables are not numeric.

##Select Predictors
First, non numeric variables need to be removed:
```{r}
nonPredictor <- 1:7
fTraining <- training[, -nonPredictor]
fTesting  <- testing[, -nonPredictor]
```

##Remove Columns with NA
Columns with NA are also eliminated:

```{r}
isNAColumn <- apply( fTraining , 2 , function(x) any(is.na(x)) )

fTraining <- fTraining[, !isNAColumn]
fTesting  <- fTesting[, !isNAColumn]
```

##Near Zero Variables
In order to improve the model, near zero variables are deleted from the data set as well:

```{r}
nzv <- nearZeroVar(fTraining)
fTraining <- fTraining[, -nzv]
fTesting <- fTesting[, -nzv]
```

##Identifying Correlated Predictors
Finally, highly correlated variables are also removed:

```{r}
highlyCor <- findCorrelation(cor(fTraining[,-ncol(fTraining)]), cutoff = 0.75)
fTraining <- fTraining[, -highlyCor]
fTesting <- fTesting[, -highlyCor]
```

##Final Training set
The resulting data set has now 41 variables:

```{r}
summary(fTraining)
```

#Model Creation
This is a classification problem. Generalized Boosted Regression Model is proposed as a method: 

```{r, eval=FALSE}
modelFitGbm <- train( classe ~.,
                      data=fTraining,
                      method="gbm",
                      verbose=FALSE)
```
```{r, echo=FALSE}
load("modelFitGbm.RData")
plot(modelFitGbm)
```

## In and Out sampling error
In order to analyse the error in the model, confusion matrices are generated for Training and Testing data set:

```{r}
confusionMatrix(fTraining$classe,predict(modelFitGbm,fTraining))
confusionMatrix(fTesting$classe,predict(modelFitGbm,fTesting))
```

